### 2017-12-27 sabs-2017/paper/bio_MTedit.tex

\textbf{Properties of a VTS that satisfies all cell-biological constraints:} Suppose we are given a particular transport graph, a particular labeling of all the compartments and edges, a particular fusion pairing matrix, and a particular regulatory model. This information is then sufficient to check the following properties, which summarise the cell-biological constraints described above:
\begin{enumerate}
\item We can determine which molecules are active on every compartment or vesicle.
\item For every vesicle fusing to a compartment, we can determine whether there exists an active pair (one molecule on the vesicle, one on the compartment) which drives that fusion event.
\item For every vesicle-compartment pair where the vesicle does not fuse to the compartment, we can verify that there is no pairing of active molecules on the
vesicle and compartment that could drive their fusion.
\item We can verify that every molecular type entering a compartment also leaves the compartment, and also that every molecular type entering a set of compartments also leaves that set. This is the steady state condition. In the biological literature this is often referred to as “homeostasis” and is a widespread and well-accepted assumption about cellular behaviour, at least over timescales of minutes to hours \cite{mani2017}.
\end{enumerate}

@@

Can we find a simple test to see whether any information is missing, given the experimental data? We have shown that graph k-connecte furnishes precisely such a test \cite{shukla}. If the data provided by cell-biologists, suitably represented as a graph, does not have a certain degree of connectivity, this implies that some biological data has been missed. (The converse is not true: even if the required connectivity does hold, there might of course be more information missing.)

>>>>

\textbf{Properties of a VTS that satisfies all cell-biological constraints:} Suppose we are given a particular transport graph, a particular labeling of all the compartments and edges, a particular fusion pairing matrix, and a particular regulatory model. This information is then sufficient to check the following properties, which summarise the cell-biological constraints described above:
\begin{enumerate}
\item We can determine which molecules are active on every compartment or vesicle.
\item For every vesicle fusing to a compartment, we can determine whether there exists an active pair (one molecule on the vesicle, one on the compartment) which drives that fusion event.
\item For every vesicle-compartment pair where the vesicle does not fuse to the compartment, we can verify that there is no pairing of active molecules on the
vesicle and compartment that could drive their fusion.
\item We can verify that every molecular type entering a compartment also leaves the compartment, and also that every molecular type entering a set of compartments also leaves that set. This is the steady state condition. In the biological literature this is often referred to as “homeostasis” and is a widespread and well-accepted assumption about cellular behaviour, at least over timescales of minutes to hours \cite{mani2016stacking}.
\end{enumerate}

@@

Can we find a simple test to see whether any information is missing, given the experimental data? We have shown that graph k-connectedness furnishes precisely such a test \cite{shukla}. If the data provided by cell-biologists, suitably represented as a graph, does not have a certain degree of connectivity, this implies that some biological data has been missed. (The converse is not true: even if the required connectivity does hold, there might of course be more information missing.)

### 2017-12-27 sabs-2017/paper/experiments.tex

To illustrate usability of our tool 
in the last column of the table~\ref{tab:grph}, we present the minimum connectedness needed for
the different variants after applying our tool for sizes from 2 to 10. 
%
We found no graph for the variant A with constraints Ann and Aen.
%
Replacing constraint on the node of Ann with Anb (variant B) does not affect the outcome.
%
If we allow every present molecule to stay active (Ann) but constraint the edge
by a boolean function (Aeb) the resultant VTS has to be at least 3-connected. Similarly, the results for the other cases are presented in the table. 
%
% If we constrain the activity of both the edge and the nodes with the boolean function (Anb and Aeb), 
% the VTS has to be 2-connected.
%
We ran experiments varying all parameters for specific N (size). For the comparison between both encodings (Table 2) we have fixed the total number of molecules to be twice as number of the nodes for N $>$ 2 and M = 2N + 1 for N = 2. For each version we also fix total number of parallel edges to maximum two. We have specifically chosen the same connectivity as Table 1 just for illustration purpose; for example in Table~\ref{tab:qf-grabh} Variant A, only check against graph with connectivity 2 is shown, in Variant B with connectivity 3 and similarly for the rest of the Variants. Choice of Variants in the table reflects different type of results variations, for example we left out version B and E as it is of same connectivity type as of A \footnote{Complete result is present in our referenced repository}.

@@

% In Table~\ref{tab:qf-grabh}, we present the running times for the search of
% VTSs of sizes 2 to 10 that satisfy the variants.
% and compare with our old encoding
% ({Old-e}) from~\cite{shukla}.
% %
% For the comparison between both encodings we have fixed the total
% number of molecules to be $|M| = 2|N|$ for $ |N|> 2$ and
% $|M| = 2|N| + 1$ for $|N| = 2$.
% %
% For each variant, we fix maximum number of parallel
% edges to 2.
% %
% In the table we have shown comparison for specific connectivity, for
% example variant A is checked against any graph with connectivity 2,
% variant B with connectivity 3 similarly for the rest of the Variants.
%
The experiments were done on a machine with Intel(R) Core(TM) i3-4030U
CPU @ 1.90GHz processor and 4GB RAM.
%
We have compared our performance with the performance of our earlier 
CBMC based implementation (old-encoding).
%
For example, the formula for variation F, the total number of
compartments ($|N|$) equals to 10, returns in 129.78 minutes (7786.8 secs)
with a SAT result.
%
In comparison, CBMC results in OUT OF MEMORY for $|N|$ greater than 5.
%
``!'' indicate that the constraints were unsatisfiable.
%
Using this encoding in comparison to the old one, not only we got efficiency improved for finding a SAT model but also did better in the case of refutation that no model exist (Table 2 Variant A timing comparison and Variant D with N =2). Hence with the use of this novel encoding, we are able to scale the system to a much larger compartmentalized system, especially to
eukaryotic cells with a total number of 10 compartments.
%
% \ankit{
% This additional scalability can provide us leverage to ask for related
% open questions that were not previously possible, for example, what is
% the minimum number of molecules required to satisfy the vesicular
% traffic system in eukaryotes.}
%
Furthermore, we experimented with limits of our tools and found
that $\zthree$ was able to solve the constraints up to $\sim{14-18}$ nodes.

>>>>

To illustrate usability of our tool 
in the last column of the table~\ref{tab:grph}, we present the minimum connectedness needed for
the different variants after applying our tool for sizes from 2 to 10. 
%
We found no graph for the variant A with constraints Ann and Aen.
%
Replacing constraint on the node of Ann with Anb (variant B) does not affect the outcome.
%
If we allow every present molecule to stay active (Ann) but constraint the edge
by a boolean function (Aeb) the resultant VTS has to be at least 3-connected. Similarly, the results for the other cases are presented in the table. 
%
% If we constrain the activity of both the edge and the nodes with the boolean function (Anb and Aeb), 
% the VTS has to be 2-connected.
%
% We ran experiments varying all parameters for specific N (size). For the comparison between both encodings (Table 2) we have fixed the total number of molecules to be twice as number of the nodes for N $>$ 2 and M = 2N + 1 for N = 2. For each version we also fix total number of parallel edges to maximum two. We have specifically chosen the same connectivity as Table 1 just for illustration purpose; for example in Table~\ref{tab:qf-grabh} Variant A, only check against graph with connectivity 2 is shown, in Variant B with connectivity 3 and similarly for the rest of the Variants. Choice of Variants in the table reflects different type of results variations, for example we left out version B and E as it is of same connectivity type as of A \footnote{Complete result is present in our referenced repository}.

@@

In Table~\ref{tab:qf-grabh}, we present the running times for the search of
VTSs of sizes 2 to 10 that satisfy the variants.
and compare with our old encoding
({Old-e}) from~\cite{shukla}.
%
For the comparison between both encodings we have fixed the total
number of molecules to be $|M| = 2|N|$ for $ |N|> 2$ and
$|M| = 2|N| + 1$ for $|N| = 2$.
%
For each variant, we fix maximum number of parallel
edges to 2.
%
In the table we have shown comparison for specific connectivity, for
example variant A is checked against any graph with connectivity 2,
variant B with connectivity 3 similarly for the rest of the Variants.

The experiments were done on a machine with Intel(R) Core(TM) i3-4030U
CPU @ 1.90GHz processor and 4GB RAM.
%
We have compared our performance with the performance of our earlier 
CBMC based implementation (old-encoding).
%
For example, the formula for variation F, the total number of
compartments ($|N|$) equals to 10, returns in 129.78 minutes (7786.8 secs)
with a SAT result.
%
In comparison, CBMC results in OUT OF MEMORY for $|N|$ greater than 5.
%
``!'' indicate that the constraints were unsatisfiable.
%
Using this encoding in comparison to the old one, not only we got efficiency improved for finding a SAT model but also did better in the case of refutation that no model exist (Table 2 Variant A timing comparison and Variant D with N =2). Hence with the use of this novel encoding, we are able to scale the system to a much larger compartmentalized system, especially to
eukaryotic cells with a total number of 10 compartments.
%
% \ankit{
% This additional scalability can provide us leverage to ask for related
% open questions that were not previously possible, for example, what is
% the minimum number of molecules required to satisfy the vesicular
% traffic system in eukaryotes.}
%
Furthermore, we experimented with limits of our tools and found
that $\zthree$ was able to solve the constraints up to $\sim{14-18}$ nodes.
